---
title: "Aides Drug"
author: "Dana Schroer"
date: "Spring 2024"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

We will use the Dyplr tool "Filter" to create a subset of the data, sorting only data where drug user = "yes". We will then use the CandP helper files to get R code to show a combination of graphical, numerical, and inferential to test the null hypothesis. 

```{r include=FALSE}

DrugUsers <- filter(triUmp, DrugUser == "yes")

# Assuming you want to reorder "treatment" based on custom levels

DrugUsers$treatment <- factor(DrugUsers$treatment, levels = c("placebo", "drug"))

```


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}

barchartGC(~infected + treatment,data=DrugUsers)
barchartGC(~infected + treatment,data=DrugUsers, type="percent")

```

    We have generated two bar charts to compare the data, the first shows total number or people who are involved in the analysis. In chart 1, we see that most people who are drug users do not contract AIDS, however it does not show us much about the effectiveness of the drug. In chart 2, we have a clearer picture because it shows the data as a percentage. In this test we see that the treatment group has lower incidence of AIDS when compares to the placebo group.


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=DrugUsers)
rowPerc(table1)
colPerc(table1)
```

    In this version of the results we see the number of people in each group and the rate at which they contracted AIDS in the infected group, we see that there are approximately three times as many cases of aids in the placebo group as opposed to the treatment group.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisqtestGC(table1)
fisherResult <- fisher.test(table1)

#Print Fisher Result
fisherResult

#Confidence Intervals
low=(fisherResult$conf.int[1]/(fisherResult$conf.int[1] + 1))*100
hi=(fisherResult$conf.int[2]/(fisherResult$conf.int[2] + 1))*100

#Print
low
hi
```

    We ran three inferential tests in this section, Two Chi-Squared tests, and a Fisher's Exact Test. In the Chi-Squared Test we are getting a p-value of 0.0000001242 which is far lower than 0.05 and a Chi-Squared ~28 which is far greater than our critical value for degrees of freedom in this case (1). Given these scores we can infer that there is a difference in the data.
    
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

    Through our graphical and numerical results, we can see that the placebo group contracts the disease at a greater frequency than the treatment group. This coupled with the tests of statistical inference show that there is a statistically significant difference in the data. We can conclude that there is as negative correlation between vaccination and disease frequency. Due to the high number of people in this study we can assume that random variations played a small role if any in the date. We can infer from these results, assuming the data was corrected accurately that there is causation in this case and the vaccine is in fact effective.